---
title: "Introducing ReLTER"
author: "Alessandro Oggioni, Micha Silver, Paulo Tagliolato"
date: "13/04/2022"
output:
  ioslides_presentation:
    smaller: true
    logo: eLTER_logo_small.png
    css: styles.css
---

```{r relter-setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is ReLTER

```{r relter-loading, message=FALSE, warning=FALSE}
# Convenient way to load list of packages
pkg_list <- c("sf", "terra", "ReLTER", "tmap")
lapply(pkg_list,require, character.only = TRUE)
tmap_options(check.and.fix = TRUE)
tmap_mode("view")
```

----

#### Package Maintainer

```{r relter-maintainer}
library(ReLTER)
maintainer("ReLTER")
```

----

#### How to cite

```{r relter-citation}
citation("ReLTER")
```

----

#### Basic functions

```{r relter-functions}
ls("package:ReLTER")
```

----

#### Categories of functions

  - site specific functions
  - network functions
  - metadata functions
  - taxonomy functions
 
## Examples

Search for DEIMS ID for a particular site. The function `get_ilter_generalinfo` allows to search by country name and site name. Then `get_site_info()` obtains various metadata about a chosen site.

#### For the first example, the Doñana LTSER Platform in Spain

```{r relter-donana}
donana = get_ilter_generalinfo(country_name = "Spain",
                              site_name = "Doñana")
(donana_id = donana$uri)
```

----

Plot a basic map of that site. We use the `tmap` package for viewing maps.

```{r relter-plot-donana}
donana_polygon <- get_site_info(donana_id, category = "Boundaries")
tm_basemap("OpenStreetMap.Mapnik") + 
  tm_shape(donana_polygon) +
  tm_fill(col = "blue", alpha = 0.3)
```
---- 

#### Next example: Retrieve metadata about a site. 

This example retrieves metadata from Lock Kinord in Scotland.

```{r relter-kinord-info}
loch_kinord <- get_ilter_generalinfo(country_name = "United K",
                              site_name = "Loch Kinord")
(loch_kinord_id = loch_kinord$uri)
loch_kinord_details <- get_site_info(loch_kinord_id,
                                 c("Contacts", "EnvCharacts", "Parameters"))

print(paste("Site manager:",
            loch_kinord_details$generalInfo.siteManager[[1]]['name'],
            loch_kinord_details$generalInfo.siteManager[[1]]['email']))
```

----

```{r relter-kinord-metadata}
# Metadata contact:
(loch_kinord_details$generalInfo.metadataProvider[[1]]['name'])
print(paste("Average air temperature:",
            loch_kinord_details$envCharacteristics.airTemperature.avg))
print(paste("Annual precipitation:",
            loch_kinord_details$envCharacteristics.precipitation.annual))
```

----

```{r relter-kinord-geobonbiome}
print(paste("GeoBonBiome:",
            loch_kinord_details$envCharacteristics.geoBonBiome[[1]]))
# Parameters:
head(loch_kinord_details$parameter[[1]]['parameterLabel'], 12)
```

----

#### Next example: Query a network

The LTER network in Slovakia

```{r elter-slovakia}
lter_slovakia_id = "https://deims.org/networks/3d6a8d72-9f86-4082-ad56-a361b4cdc8a0"

network_research_topics <- get_network_research_topics(lter_slovakia_id)
head(network_research_topics$researchTopicsLabel, 20)
```

----

```{r relter-slovakia-network}
lter_slovakia_sites <- get_network_sites(lter_slovakia_id)
lter_slovakia_sites$title
```

----

#### Show map of sites in the network

```{r relter-slovakia-map}
lter_slovakia <- produce_network_points_map(lter_slovakia_id, "SVK")
svk <- readRDS("gadm36_SVK_0_sp.rds")  # downloaded by produce_network_points_map()
tm_basemap("OpenStreetMap.Mapnik") + 
  tm_shape(lter_slovakia) + 
  tm_dots(col = "blue", size=0.04) +
  tm_shape(svk) + 
  tm_borders(col = "purple", lwd = 0.6) +
  tm_grid(alpha = 0.4) +
  tm_scale_bar(position = c("right", "bottom"))
```

